---
title: "Covid91 vaccine study FinalFall2022"
author: "Wesley Cruz"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalFall2022.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

```{r}
Males_data <- filter(FinalData, sex == "M")
```

```{r}
barchartGC(~infected + treatment,data=Males_data)
barchartGC(~infected + treatment,data=Males_data , type="percent", main = ("Infected vs Treatment"))
```

The barplots displayed above depict male cases from clinical trials. Blue bars represent males who received a placebo or imitation medication, while pink bars signify those who were administered a vaccine or potent medication. The first bar chart, labeled with 'frequency' on the y-axis, conveys the number of individuals in specific categories within the entire dataset. Meanwhile, the second chart, marked with 'percent' on the y-axis, illustrates the effectiveness rate of these treatments over a recurring time frame or within a given sample.

In this particular dataset, blue bars are situated above the pink bars at roughly 70% and 50%. Conversely, pink bars are either below or on par with the blue bars, hovering around 30% and 50%.

### Numerical results
```{r}
table1 <- xtabs(~infected + treatment, data=Males_data)
rowPerc(table1)
colPerc(table1)
```

The numerical data for males presented in the tables show the statistical outcomes for individuals who are infected, healthy, and either vaccinated or not vaccinated. In the first table, out of 100 males, 68 who are ill receive a placebo or imitation medication, while 32 are given a vaccine. The second table outlines the results from the trial.

The likelihood of a positive COVID-19 test for those receiving a placebo is 2.7, while for those receiving the vaccine, the probability drops to 1.2. As for patients with negative test results, the percentages for those under placebo and vaccine conditions are 97.3 and 98.8, respectively. Based on this data, men who have been vaccinated are at a lower risk compared to those who have not received the vaccine.

### Inferential

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
malesfish <- fisher.test(table1)
malesfish
```

The primary focus of our analysis is the Fisher's Exact test. In this case, the test results show a p-value of 2e-12, an alternative hypothesis stating that the true odds ratio is not equal to 1, a 95% confidence interval ranging from 1.7 to 2.8, and sample estimates with an odds ratio of 2.2. Based on these findings, we can conclude that there is no association between the treatment's effectiveness across the monitored subpopulation, as the p-value is smaller than the significance threshold of 0.05.

## Effectiveness Confidence Interval 'r low' to 'r hi'

```{r}
effect <- malesfish$p.value/(malesfish$p.value +1)
effectivenessmales = malesfish$estimate/(malesfish$estimate +1)*100
print(effectivenessmales)
```

```{r}
low=(malesfish$conf.int[1]/(malesfish$conf.int[1] +1))*100
hi=(malesfish$conf.int[2]/(malesfish$conf.int[2] +1))*100
options(digits=2)
```

The odds ratio represents the efficacy of the vaccine administered to a group of males based on 18,125 observations. While the probability of the vaccine being effective for males is 69, there isn't sufficient evidence to support this outcome. This conclusion is drawn by considering the earlier barplot, numerical results, and the inferential Fisher's Exact Test for Count Data.

## Females
```{r}
females_data <- filter(FinalData, sex == "F")
```

```{r}
barchartGC(~infected + treatment,data=females_data)
barchartGC(~infected + treatment,data=females_data , type="percent", main = ("Infected vs Treatment"))
```

The barplots above showcase female cases from clinical trials. Although the outcomes for females differ from those for males, there are still distinct groups of females who are either sick or healthy. In these barplots, blue bars represent females who received a placebo or imitation medication, while pink bars signify females who were administered a vaccine or potent medication. On the left, the chart shows those who still contracted COVID-19 despite the treatment, while on the right, it displays those who were successfully protected against the virus.

### Numerical results
```{r}
table1 <- xtabs(~infected + treatment, data=females_data)
rowPerc(table1)
colPerc(table1)
```

The numerical data for females present the statistical outcomes for infected (vaccinated), infected (unvaccinated), healthy (vaccinated), and healthy (unvaccinated) individuals. In the first table, among 100 sick females, 43 experienced a placebo effect, and 57 received a vaccine. In the same group, 50 placebo and 50 vaccine females were deemed healthy. The second table outlines the trial results.

The probability of a positive COVID-19 test for those undergoing a placebo treatment is 1.7, while for those receiving the vaccine, the likelihood increases to 2.3. As for patients with negative test results, the percentages for those under placebo and vaccine conditions are 98.3 and 97.7, respectively. Interestingly, in this specific dataset, more females were protected after experiencing a placebo effect, making the vaccine appear less valuable compared to the placebo.


### Inferential 
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
femalesfish <- fisher.test(table1)
femalesfish
```

Similar to the male results, we can analyze the Fisher's Exact test to gain insight into the effectiveness of treatments across the female subpopulation. The test results indicate a p-value of 0.005, an alternative hypothesis stating that the true odds ratio is not equal to 1, and a 95% confidence interval ranging from 0.60 to 0.92. The sample estimates indicate an odds ratio of 0.74. Based on these results, we can assume that there is also no association between the effectiveness of treatments across the female subpopulation that was monitored.


## Effectiveness Confidence Interval 'r low' to 'r hi'
```{r}
effect <- femalesfish$p.value/(femalesfish$p.value +1)
effectivenessfemales = femalesfish$estimate/(femalesfish$estimate +1)*100
print(effectivenessfemales)
```

```{r}
low=(femalesfish$conf.int[1]/(femalesfish$conf.int[1] +1))*100
hi=(femalesfish$conf.int[2]/(femalesfish$conf.int[2] +1))*100
options(digits=2)
```


Typically, an odds ratio indicates a strong association, but since the odds ratio for females is smaller than that for males, there is a possibility that the vaccine may not be as effective for females. Thus, it seems that there may not be a significant association between the effectiveness of the treatment across the COVID-19 virus for females. For example, the analysis of female results has demonstrated that there is no clear association in effectively preventing COVID-19 for females.

## LGBTQ

```{r}
LGBTQ_data <- filter(FinalData, LGBTQ == "gay")
```

```{r}
barchartGC(~infected + treatment,data=LGBTQ_data)
barchartGC(~infected + treatment,data=LGBTQ_data , type="percent", main = ("Infected vs Treatment"))
```


The above barplots illustrate the percentage of LGBTQ individuals who were infected with COVID-19 and those who were healthy. Pink bars represent those who received the vaccine, while blue bars signify those who received the placebo. Notably, there was a significantly higher number of COVID-19 patients who received the vaccine than those who received the placebo. Additionally, there were slightly more patients who were healthy and received the vaccine compared to those who received the placebo.

### Numerical results
```{r}
table1 <- xtabs(~infected + treatment, data=LGBTQ_data)
rowPerc(table1)
colPerc(table1)
```

The numerical data for LGBTQ individuals presents the statistical outcomes for infected, healthy, and vaccinated or unvaccinated individuals. In the first table, out of 100 individuals who were sick, 13 experienced a placebo effect, while 87 were given a vaccine. Among the same group, 49 individuals received a placebo, while 51 individuals received the vaccine and were deemed healthy. The second table outlines the results for COVID-19 and healthy individuals for those who received the placebo and those who received the vaccine.

According to this data, individuals were not fully protected as there was still a 3.3 chance of contracting COVID-19 even after receiving the vaccine. Specifically, the probability of contracting COVID-19 after receiving the placebo was 0.54, whereas the chance for the vaccine was 3.3. Additionally, the percentage of healthy individuals after receiving the placebo was 99.46, while for the vaccine, the rate was 96.7.

### Inferential 
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
LGBTQfish <- fisher.test(table1)
LGBTQfish
```   

The results of the Fisher's Exact test in the case of LGBTQ individuals show a p-value of 1e-08, an alternative hypothesis stating that the true odds ratio is not equal to 1, and a 95% confidence interval ranging from 0.066 to 0.344. The sample estimates indicate an odds ratio of 0.16. Therefore, we can also assume that there is no clear association between the effectiveness of the vaccine across the variety of subpopulations that were monitored.

Furthermore, the chance that the vaccine was effective in protecting LGBTQ individuals from COVID-19 is slightly lower than the probability for the vaccine's effectiveness in protecting males from COVID-19, with the LGBTQ value at 13.9 and the males value at 17.2. Based on these findings, the vaccine may not be effective in future clinical trials for both LGBTQ and male individuals.

## Druggies


# Overall Results and Conclusions