---
title: "Covid91 vaccine study FinalFall2022"
author: "Wesley Cruz"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalFall2022.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

```{r}
Males_data <- filter(FinalData, sex == "M")
```

```{r}
barchartGC(~infected + treatment,data=Males_data)
barchartGC(~infected + treatment,data=Males_data , type="percent", main = ("Infected vs Treatment"))
```

The barplots displayed above depict male cases from clinical trials. Blue bars represent males who received a placebo or imitation medication, while pink bars signify those who were administered a vaccine or potent medication. The first bar chart, labeled with 'frequency' on the y-axis, conveys the number of individuals in specific categories within the entire dataset. Meanwhile, the second chart, marked with 'percent' on the y-axis, illustrates the effectiveness rate of these treatments over a recurring time frame or within a given sample.

In this particular dataset, blue bars are situated above the pink bars at roughly 70% and 50%. Conversely, pink bars are either below or on par with the blue bars, hovering around 30% and 50%.

### Numerical results
```{r}
table1 <- xtabs(~infected + treatment, data=Males_data)
rowPerc(table1)
colPerc(table1)
```

The numerical data for males presented in the tables show the statistical outcomes for individuals who are infected, healthy, and either vaccinated or not vaccinated. In the first table, out of 100 males, 68 who are ill receive a placebo or imitation medication, while 32 are given a vaccine. The second table outlines the results from the trial.

The likelihood of a positive COVID-19 test for those receiving a placebo is 2.7, while for those receiving the vaccine, the probability drops to 1.2. As for patients with negative test results, the percentages for those under placebo and vaccine conditions are 97.3 and 98.8, respectively. Based on this data, men who have been vaccinated are at a lower risk compared to those who have not received the vaccine.

### Inferential

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
malesfish <- fisher.test(table1)
malesfish
```

The primary focus of our analysis is the Fisher's Exact test. In this case, the test results show a p-value of 2e-12, an alternative hypothesis stating that the true odds ratio is not equal to 1, a 95% confidence interval ranging from 1.7 to 2.8, and sample estimates with an odds ratio of 2.2. Based on these findings, we can conclude that there is no association between the treatment's effectiveness across the monitored subpopulation, as the p-value is smaller than the significance threshold of 0.05.



## Females


## LGBTQ



## Druggies


# Overall Results and Conclusions