---
title: "Covid91 vaccine study FinalFall2022"
author: "Wesley Cruz"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalFall2022.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

```{r}
Males_data <- filter(FinalData, sex == "M")
```

```{r}
barchartGC(~infected + treatment,data=Males_data)
barchartGC(~infected + treatment,data=Males_data , type="percent", main = ("Infected vs Treatment"))
```

The barplots displayed above depict male cases from clinical trials. Blue bars represent males who received a placebo or imitation medication, while pink bars signify those who were administered a vaccine or potent medication. The first bar chart, labeled with 'frequency' on the y-axis, conveys the number of individuals in specific categories within the entire dataset. Meanwhile, the second chart, marked with 'percent' on the y-axis, illustrates the effectiveness rate of these treatments over a recurring time frame or within a given sample.

In this particular dataset, blue bars are situated above the pink bars at roughly 70% and 50%. Conversely, pink bars are either below or on par with the blue bars, hovering around 30% and 50%.


## Females


## LGBTQ



## Druggies


# Overall Results and Conclusions